Efficient inference using the approximate Riemannian geometry of stochastic population models

### 2012-08-07 paper/paper.tex
Copied material on the immigration death process from the documents in docs to the appendix.

\author{B. Calderhead\affil{1}{UCL}, C. S. Gillespie\affil{2}{NCL}}
% \email{b.calderhead@ucl.ac.uk}
% \email{colin.gillespie@newcastle.ac.uk}
% \\
% 	   Centre for Mathematics and Physics in the Life Sciences and Experimental Biology,\\University College London, UK
% 	   \and 
% 	   School of Mathematics \& Statistics, Newcastle University,
% 	   NE1 7RU, UK
\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

@@

%% When adding keywords, separate each term with a straight line: |
\keywords{term | term | term}

@@

%% Optional Appendix or Appendices
%% \appendix Appendix text...
%% or, for appendix with title, use square brackets:
%% \appendix[Appendix Title]

>>>>

\author{Ben Calderhead\affil{1}{University College London, London, United Kingdom}, Colin S. Gillespie\affil{2}{Newcastle University, Newcastle upon Tyne, United Kingom}}
% \email{b.calderhead@ucl.ac.uk}
% \email{colin.gillespie@newcastle.ac.uk}
% \\
% 	   Centre for Mathematics and Physics in the Life Sciences and Experimental Biology,\\University College London, UK
% 	   \and 
% 	   School of Mathematics \& Statistics, Newcastle University,
% 	   NE1 7RU, UK
\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

@@

%% When adding keywords, separate each term with a straight line: |
\keywords{moment closure | stochastic | inference}

@@

%% Optional Appendix or Appendices
%% \appendix Appendix text...
%% or, for appendix with title, use square brackets:

%Taken from the tex file in the docs folder
%Not edited yet
\appendix[Immigration-death process]

This model has a single species, $N(t)$. The first reaction represents a constant immigration rate, and occurs with rate $k_1$. The second reaction represents death in the $N$ population and occurs with rate $k_2 N(t)$. The model can be represented by the coupled (pseudo) reactions, 
\begin{equation}\label{1}
R_1: \emptyset  \xrightarrow{\phantom{a}k_1\phantom{a}} N 
\quad \text{and} \quad
R_2: N  \xrightarrow{\phantom{a}k_2\phantom{a}} \emptyset.
\end{equation}
The forward Kolmogorov equation (or CME) for this model is
\[
dp_{n}(t)/dt = k_1 p_{n-1}(t) +  k_2 (n+1) p_{n+1}(t) 
- (k_1 + k_2 n) p_{n}(t) \;.
\]
This equation can be solved exactly using a generating function approach.

\subsection{Moments}

The general idea of \textit{moment closure} is to represent the analytically intractable moments of the process as a set of ODEs. The solutions of these ODEs provide approximations for the moments. With this particular system, we can solve the ODEs analytically - so the moments are exact.

The marginal means of the process are:
\[
E[n(t) | n(0) = n_0 ] = \frac{k_1}{k_2}(1-e^{-k_2 t}) + n_0 e^{-k_2 t} \;.
\]
Notice that as $t \rightarrow \infty$,
\[
E[n(t) | n(0) = n_0 ] = \frac{k_1}{k_2} \;.
\]
The marginal variance is
\[
Var[n(t) | n_0]  = \left(\frac{k_1}{k_2} + n_0e^{-k_2 t}\right)(1-e^{-k_2 t}) \;.
\]
